---
title: "Analysis of First Pass Regressions Out "
author: "Henrik Singmann"
output: 
  html_document:
    toc: yes
    toc_float: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


```{r setup, results='hide', message=FALSE}
require(dplyr)
require(tidyr)
require(stringr)
require(ggplot2)
require(afex)
#require(lattice) # for qqmath
require(gridExtra)
theme_set(theme_light())
require(optimx)
require(DHARMa) # for residual analysis
#require(binom)
#require(effects)
```

# Introduction

The goal of the following analysis was to see whether the explanatory goodness of fit would affect the DVs. For detailed analysis of the IVs see other file.

```{r}
data_in <- read.csv("RIs.csv")
data_in$P.s <- factor(data_in$P.s)
data_in$Item <- factor(data_in$Item)
```


# Regressions In

## Data Preparation


```{r add_rating}
RPs_plus_ratings <- read.csv("RPs_plus_ratings.csv")
RPs_plus_ratings$P.s <- factor(RPs_plus_ratings$P.s)
RPs_plus_ratings$Item <- factor(RPs_plus_ratings$Item)

agg1 <- RPs_plus_ratings %>% 
  group_by(Item, Fit) %>% 
  summarise(Accepts = first(Accepts),
            Confident = first(Confident))
agg2 <- RPs_plus_ratings %>% 
  group_by(Item, Fit) %>% 
  summarise(Accepts = mean(Accepts),
            Confident = mean(Confident))
stopifnot(all.equal(agg1, agg2)) # to make sure nothing fishy is going on.

data_in <- left_join(data_in, agg1, by = c("Item", "Fit"))
```


```{r}
drp_r <- droplevels(data_in[data_in$Item != "9" ,])
nas <- is.na(drp_r$spillover) | is.na(drp_r$Consequent)
drp_r <- droplevels(drp_r[!nas,])

```

For the analysis we remove `r sum(nas)` trials with missing data in either spillover or Consequent. Then we z-transform the continuous `Accepts` variable.

```{r}
drp_r <- drp_r %>%
  mutate(
    acc_z = (Accepts - mean(Accepts))/sd(Accepts))
    #conf_z = (Confident- mean(Confident))/sd(Confident))
```


# Consequent

Note that the maximal model for models including `Fit` have random slopes for `Fit` for both random effects grouping factors (i.e., participant `P.s` and `Item`). For models including the continuous fit measure `Accepts` we have by-participant random slopes for `Accepts`, but by-item random slopes for `Fit` (as `Accepts` has only two different values for each `Item`,  using the binary `Fit` variable seems more reasonable.)


```{r, fig.width=6, fig.height=4}

agg <- drp_r %>% 
  group_by(P.s, Fit) %>% 
  summarise(prob_consequent = mean(Consequent),
            prob_spillover = mean(spillover),
            consequent = sum(Consequent),
            spillover = sum(spillover),
            n = n()) %>% 
  ungroup 

p_obs_acc <- ggplot(data = drp_r, aes(x = acc_z, y = Consequent)) +
  geom_point()

p_obs_fit <- ggplot(agg, aes(x = Fit, y = prob_consequent)) +
  geom_boxplot() +
   stat_summary(fun.y = "mean", colour = "red", size = 2, geom = "point")
grid.arrange(p_obs_fit, p_obs_acc, ncol = 2)

```


## With Fit

```{r}
m_con_f_1 <- mixed(Consequent ~ Fit + (Fit|P.s) + (Fit| Item), drp_r, family = binomial, method = "LRT", progress = FALSE)
summary(m_con_f_1)
```

The correlations between the random effects parameters are at the boundary of the parameter space, consequently we remove those.

```{r}
m_con_f_2 <- mixed(Consequent ~ Fit + (Fit||P.s) + (Fit|| Item), drp_r, family = binomial, method = "LRT", progress = FALSE, expand_re = TRUE)
summary(m_con_f_2)
```
Now the by-participant and by-item random slopes for `Fit` are virtually zero, so we consequently refit the model without these parameters, we also try all available optimizer to see if this helps.

```{r}
m_con_f_3 <- mixed(Consequent ~ Fit + (1|P.s) + (1| Item), drp_r, family = binomial, method = "LRT", progress = FALSE, expand_re = FALSE, all_fit = TRUE)
summary(m_con_f_3)
```

Overall, the final model seems to provide the best account (i.e., is the optimal model).

```{r}
anova(m_con_f_1, m_con_f_2, m_con_f_3)
```


We also perform some graphical model fit tests (using `DHARMa`, see: https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html) which suggest the model is without large problems (i.e., for the QQ-plot the residuals are on the main diagonal and for the residuals versus predicted the red lines are straight and at .25, .5 and .75).

```{r, fig.width=8, fig.height=4}
sim_m_con_f_3 <- simulateResiduals(fittedModel = m_con_f_3$full_model, n = 250)
plotSimulatedResiduals(sim_m_con_f_3)
```

Interestingly, only for the optimal model but not for the maximal model is the effetc of Fit significant.

```{r}
nice(m_con_f_1)
nice(m_con_f_3)
```

## With Accepts (non-transformed)

```{r}
m_con_a_1 <- mixed(Consequent ~ acc_z + (acc_z|P.s) + (Fit| Item), drp_r, family = binomial, method = "LRT", progress = FALSE)
summary(m_con_a_1)
```

We see a suspcisouly high correlations between for both random effect grouping factors, consequently we remove those.


```{r}
m_con_a_2 <- mixed(Consequent ~ acc_z + (acc_z||P.s) + (Fit|| Item), drp_r, family = binomial, method = "LRT", progress = FALSE, expand_re = TRUE, all_fit = TRUE)
summary(m_con_a_2)
```

Given the small estimate of the by-participant random slope we remove it.

```{r}
m_con_a_3 <- mixed(Consequent ~ acc_z + (1|P.s) + (1| Item), drp_r, family = binomial, method = "LRT", progress = FALSE, expand_re = TRUE, all_fit = TRUE)
summary(m_con_a_3)
```

```{r}
anova(m_con_a_1, m_con_a_2, m_con_a_3)
```

Again, the reduced model seems to be the optimal model. Consequently we also inspect the residuals of this model. Overall this looks very reasonable despite the fitted versus residuals plot.

```{r, fig.width=8, fig.height=4}
sim_m_con_a_3 <- simulateResiduals(fittedModel = m_con_a_3$full_model, n = 500)
plotSimulatedResiduals(sim_m_con_a_3)
```

Interestingly, neither model shows a significant effect of `Accepts` (z-transformed to `acc_z`). 

```{r}
nice(m_con_a_1)
nice(m_con_a_3)
```

This contrasts somewhat with the results of the Wald tests reported above which clearly show a significant effect of `acc_z`. Consequently, we repeated the analysis using parametric bootstrap. This also showed no effect of `acc_z`.


# Spillover

```{r, fig.width=6, fig.height=4}

p_obs_acc <- ggplot(data = drp_r, aes(x = acc_z, y = spillover)) +
  geom_point()

p_obs_fit <- ggplot(agg, aes(x = Fit, y = prob_spillover)) +
  geom_boxplot() +
   stat_summary(fun.y = "mean", colour = "red", size = 2, geom = "point")
grid.arrange(p_obs_fit, p_obs_acc, ncol = 2)

```

Given there is only one spillover event this obviates all statistical analyses.
